### 2017-10-11 EUCBV/Conclusion.tex
Latest on 11/10/2017

In this paper, we studied the EUCBV algorithm which takes into account the empirical variance of the arms and  employs aggressive exploration parameters in conjunction with non-uniform arm selection (as opposed to UCB-Improved) to eliminate sub-optimal arms. Our theoretical analysis conclusively established that EUCBV exhibits an order-optimal gap-independent regret bound of $O\left(\sqrt{KT}\right)$. Empirically, we show that EUCBV performs superbly across diverse experimental settings and outperforms most of the bandit algorithms in a  stochastic MAB setup. Our experiments show that EUCBV is extremely stable for larger horizons and performs consistently well across different types of distributions. One avenue for future work is to remove the constraint of $T\geq K^{2.4}$ required for EUCBV to reach the order optimal regret bound. Another future direction is to come up with an anytime version of EUCBV. An anytime algorithm does not need the horizon $T$ as an input parameter.

%\section*{Acknowledgement} This work was supported by a funding from IIT Madras under project CSE/14-15/831/ RFTP/BRAV.