### 2017-10-11 EUCBV/EUCBV.tex

\def\year{2018}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required

\usepackage{macros}

\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (2018 Formatting Instructions for Authors Using LaTeX)
/Author (AAAI Press Staff)}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Efficient-UCBV: An Almost Optimal Algorithm using Variance Estimates}
\author{Author names withheld}

\maketitle

\begin{abstract}
We propose a novel variant of the UCB algorithm (referred to as Efficient-UCB-Variance (EUCBV)) for minimizing cumulative regret in the stochastic multi-armed bandit (MAB) setting. EUCBV incorporates the arm elimination strategy proposed in UCB-Improved \citep{auer2010ucb}, while taking into account the variance estimates to compute the arms' confidence bounds, similar to UCBV \citep{audibert2009exploration}. Through a theoretical analysis we establish that EUCBV incurs a \emph{gap-dependent} regret bound of {\scriptsize $O\left( \dfrac{K\sigma^2_{\max} \log (T\Delta^2 /K)}{\Delta}\right)$} after $T$ trials, where $\Delta$ is the minimal gap between optimal and sub-optimal arms; the above bound is an improvement over that of existing state-of-the-art UCB algorithms (such as UCB1, UCB-Improved, UCBV,  MOSS). Further, EUCBV incurs a \emph{gap-independent} regret bound of {\scriptsize $O\left(\sqrt{KT}\right)$}  which is an improvement over that of UCB1, UCBV and UCB-Improved, while being comparable with that of MOSS and OCUCB. Through an extensive numerical study we show that EUCBV significantly outperforms the popular UCB variants (like MOSS, OCUCB, etc.) as well as Thompson sampling and Bayes-UCB algorithms.

%\Keywords{Stochastic multi-armed bandits, cumulative regret, UCB-Improved, UCBV.}
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

\subsection{Related Work}
\label{sec:related}
\input{related}

\subsection{Our Contributions}
\label{sec:contri}
\input{contribution}

\section{Algorithm: Efficient UCB Variance}
\label{sec:eucbv}
\input{ealgo}

\section{Main Results} 
\label{sec:results}
\input{results}

\section{Proofs}
\label{sec:proofTheorem}
\input{proofTheorem}

\section{Experiments}
\label{sec:expt}
\input{expts}

\section{Conclusion and Future Works}
\label{sec:conc}
\input{Conclusion}

\clearpage
\newpage

\bibliography{biblio}
\bibliographystyle{aaai}

\clearpage
\newpage

\section{Appendix}
\label{sec:app}
\input{Appendix}

\end{document}