### 2017-10-17 Appendix/AppendixConc.tex
Latest on 17/10/2017

\subsection{Sample space}

\subsection{Events}

\subsection{Sigma-algebra}

\subsubsection{Borel-sigma algebra}

\subsection{Measure}

\subsubsection{The probability measure}

\subsection{The triplet}

\subsection{Filtration}

\section{Martingale}

\subsection{Super-martingale}

\subsection{Sub-martingale}

\section{Convergence theorems}

\subsection{Monotone convergence theorem}

\subsection{Dominated convergence theorem}

\subsection{Fatou's Lemma}

\section{Sub-Gaussian distribution}

Let a random variable $X\in \R$ with variance as $\sigma^2$. Then $X$ is said to be $\sigma$-sub-gaussian for $\sigma\geq 0$ such that $\E[X] = 0$ and its moment generating function satisfies for all $\lambda \in \R $ the following condition,

\begin{align*}
\E[\exp{\lambda X}]\leq \exp\left( - \dfrac{\lambda^2\sigma^2}{2}\right)
\end{align*}

Also, note that sub-gaussian distribution is a class of distribution rather than a distribution itself.

\begin{remark}
A random variable $X\in[0,1]$ is said to be $\dfrac{1}{2}-sub-gaussian$ with its moment generating function satisfying the condition,

\begin{align*}
\E[\exp{\lambda X}]\leq \exp\left( - \dfrac{\lambda^2}{8}\right), \forall \lambda \in \R
\end{align*}

\end{remark}

\section{Concentration Inequalities}

In this section we state some of the concentration inequalities used in the proofs in several chapters of the thesis. Concentration inequality deals with the control of the tail of the average of independent random variables from their expected mean.

	Let, $X_1,X_2,\ldots,X_n$ be a sequence of independent random variables defined on a probability space $(\omega,\mathcal{F},\Pb)$, is bounded in $[a_i,b_i],\forall i=1,2,\ldots, n$. Let $S_n$ denote the sum of the random variables such that $S_n = X_1 + X_2 + \ldots + X_n$,  $\hat{r} = \dfrac{S_n}{n}$ and $E[S_n]=r$. Let $\mathcal{F}_n$ be an increasing sequence of $\sigma$-fields of $\mathcal{F}$ such that for each $n$, $\sigma(X_{1},\ldots,X_n)\subset \mathcal{F}_t$ and for $q>t$, $X_q$ is independent of $\mathcal{F}_n$.

\subsection{Markov's inequality}

Markov's inequality states that, for any $\epsilon > 0$,

\begin{align*}
\Pb[S_n > \epsilon] \leq \dfrac{\E[S_n]}{\epsilon}
\end{align*}

\subsection{Chernoff-Hoeffding Bound}

Chernoff-Hoeffding gives us the following inequality regarding the sums of independent random variables $S_n$ and their deviation from their expectation $\E[S_n]=r$, for any $\epsilon > 0$,

\begin{align*}
&\Pb\lbrace S_n - n\E[S_n] \geq \epsilon \rbrace \leq \exp\left( -\dfrac{2\epsilon^2}{n \sum_{i=1}^n(a_i -b_i)}\right) \\
%%%%%%%%%%%%%%%%%%%%%%
&=\Pb\lbrace S_n - n\E[S_n] \leq - \epsilon \rbrace \leq \exp\left( -\dfrac{2\epsilon^2}{n \sum_{i=1}^n(a_i -b_i)}\right)
\end{align*}

Considering all the random variables bounded in $[0,1]$, the above two inequalities can be reduced to,

\begin{align*}
&\Pb\lbrace \left|\dfrac{S_n}{n} - \E[S_n]\right| \geq \epsilon \rbrace \leq 2\exp\left( - 2\epsilon^2 n \right) \\
%%%%%%%%%%%%%%%%%%%%%%
&=\Pb\lbrace \left|\hat{r} - r\right| \geq \epsilon \rbrace \leq 2\exp\left( - 2\epsilon^2 n \right)
\end{align*}

\subsection{Empirical Bernstein inequality}

Similar to Chernoff-Hoeffding bound, empirical Bernstein inequality gives us the following inequality regarding the sums of independent random variables $S_n$ and their deviation from their expectation $\E[S_n]=r$, for any $\epsilon > 0$,

\begin{align*}
&\Pb\lbrace S_n - n\E[S_n] \geq \epsilon \rbrace \leq \exp\left( -\dfrac{2\epsilon^2}{\left(2\sigma^2 + \frac{2 b_{\max} \epsilon}{3}\right) n \sum_{i=1}^n(a_i -b_i)}\right), \\
%%%%%%%%%%%%%%%%%%%%%%
&\Pb\lbrace S_n - n\E[S_n] \leq - \epsilon \rbrace \leq \exp\left( -\dfrac{2\epsilon^2}{\left(2\sigma^2 + \frac{2 b_{\max} \epsilon}{3}\right) n \sum_{i=1}^n(a_i -b_i)}\right)
\end{align*}

Considering all the random variables bounded in $[0,1]$, the above two inequalities can be reduced to,

\begin{align*}
&\Pb\lbrace \left|\dfrac{S_n}{n} - \E[S_n]\right| \geq \epsilon \rbrace \leq 2\exp\left( -\dfrac{2\epsilon^2 n}{\left(2\sigma^2 + \frac{2\epsilon}{3}\right)}\right) \\
%%%%%%%%%%%%%%%%%%%%%%
&=\Pb\lbrace \left|\hat{r} - r\right| \geq \epsilon \rbrace \leq 2\exp\left( -\dfrac{2\epsilon^2 n}{\left(2\sigma^2 + \frac{2\epsilon}{3}\right)}\right)
\end{align*}