### 2017-10-17 Chapter4/prevResAPT.tex

\begin{algorithm}[!h]
\caption{APT}
\label{alg:apt}
\begin{algorithmic}
\State {\bf Input:} Time horizon $T$, threshold $\tau$, tolerance factor $\epsilon\geq 0$
\State Pull each arm once
\State \For{$t=K+1,..,T$}
\State Pull arm $j\in\argmin_{i\in A}\big\lbrace \left(|\hat{r}_{i} - \tau | + \epsilon\right)\sqrt{n_i}\big\rbrace$ and observe the reward for arm $j$.
\EndFor
\State \textbf{Output:} $\hat{S}_{\tau}=\lbrace i: \hat{r}_{i}\geq \tau \rbrace$.
\end{algorithmic}
\end{algorithm}

In the latest work of \citet{locatelli2016optimal} Anytime Parameter-Free Thresholding (APT) algorithm comes up with an improved anytime guarantee than CSAR for the thresholding bandit problem.