### 2017-11-04 Chapter2/conclusion.tex
Latest on 5/11/2017

In this chapter we looked at the stochastic multi-armed bandit (SMAB) setting and discussed how it is important in the general reinforcement learning setup. We also looked at the various state-of-the-art algorithms in the literature for the SMAB setting and discussed the advantages and disadvantages of them. The regret bounds that have been proven for the said algorithms have also been discussed at length and their confidence intervals have also been compared against each other. In the next chapter we provide our solution to this SMAB setting which achieves an almost order-optimal regret bound.