### 2017-11-04 Chapter2/motivation.tex

There has been a significant amount of research in the area of stochastic MABs. One of the earliest work can be traced to \citet{thompson1933likelihood}, which deals with  the problem of choosing between two treatments to administer on patients who come in sequentially. In \citet{thompson1935theory} this work was extended to include more general cases of finitely many treatments. In recent years the SMAB setting has garnered extensive popularity because of its simple learning  model and its practical applications in a wide-range of industries, including, but not limited to, mobile channel allocations, online advertising and computer simulation games. Some of these problems have been already discussed in chapter \ref{chap:intro}, section \ref{motivation} and an interested reader can refer to it.