### 2017-11-05 Chapter4/conc.tex

In this chapter, we looked at the pure exploration MAB and thresholding bandit (TBP) setting which is a special case of combinatorial pure exploration MAB. We then looked at the various state-of-the-art algorithms in the literature for the pure-exploration setting and discussed the advantages and disadvantages of them. Then we looked at the latest algorithm for the TBP setting. The expected loss that has been proven for the said algorithms have also been discussed at length and their exploration parameters have also been compared against each other. In the next chapter, we provide our solution to this TBP setting which uses variance estimation to find the set of arms above the threshold.