### 2017-11-24 Chapter6/expt.tex

In this section we present two experiments in two different environments.

\begin{figure}[!th]
    \centering
    \begin{tabular}{cc}
    %\setlength{\tabcolsep}{0.1pt}
    \subfigure[\Large\textwidth][\large Expt-$1$: $3$ Bernoulli-distributed arms (From Dr. Odalric's Draft).]
    %with $r_{i_{{i}\neq {*}}}=0.07$ and $r^{*}=0.1$
    {
    		\pgfplotsset{
		tick label style={font=\normalsize},
		label style={font=\normalsize},
		legend style={font=\normalsize},
		ylabel style={yshift=12pt},
		%legend style={legendshift=32pt},
		}
        \begin{tikzpicture}[scale=0.7]
      	\begin{axis}[
		xlabel={timestep},
		ylabel={Cumulative Regret},
		grid=major,
        %clip mode=individual,grid,grid style={gray!30},
        clip=true,
        %clip mode=individual,grid,grid style={gray!30},
  		legend style={at={(0.5,1.4)},anchor=north, legend columns=3} ]
      	% UCB

		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_DUCB01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS02.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS03.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS1E01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS1E02.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_TS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_OTS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_DTS01.txt};

      	\legend{DUCB($\gamma=1-\frac{1}{4\sqrt{T}}$),ETSDAE1,ETSDAE2,ETSDAE3,ETSD1E1,ETSD1E2,TS,OTS,DTS($\gamma=1-\frac{1}{4\sqrt{T}})$}    
      	\end{axis}
      	\end{tikzpicture}
  		\label{psbandit:fig:1}
    }
    &
    \subfigure[\Large\textwidth][\large Expt-$1$: $3$ Bernoulli-distributed arms (From Dr. Odalric's Draft).]
    %with $r_{i_{{i}\neq {*}}}=0.07$ and $r^{*}=0.1$
    {
    		\pgfplotsset{
		tick label style={font=\normalsize},
		label style={font=\normalsize},
		legend style={font=\normalsize},
		ylabel style={yshift=12pt},
		%legend style={legendshift=32pt},
		}
        \begin{tikzpicture}[scale=0.7]
      	\begin{axis}[
		xlabel={timestep},
		ylabel={Cumulative Regret},
		grid=major,
        %clip mode=individual,grid,grid style={gray!30},
        clip=true,
        %clip mode=individual,grid,grid style={gray!30},
  		legend style={at={(0.5,1.4)},anchor=north, legend columns=3} ]
      	% UCB

		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_DUCB01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS04.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS1E03.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS06.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS1E04.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_TS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_OTS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_DTS01.txt};
		\addplot table{Chapter6/results/NewExpt/Expt5/comp_subsampled_ETS1E06.txt};

      	\legend{DUCB($\gamma=1-\frac{1}{4\sqrt{T}}$),EAggrCPD1,CPD1-E1,EAggrCPD2,CPD2-E1,TS,OTS,DTS($\gamma=1-\frac{1}{4\sqrt{T}})$,CPD2-E1-N}    
      	\end{axis}
      	\end{tikzpicture}
  		\label{psbandit:fig:2}
    }
    \end{tabular}
    \caption{Cumulative regret for various bandit algorithms on a piecewise stochastic 3-armed bandit environment. }
    \label{fig:karmed1}
\end{figure}

%\begin{figure}[!th]
%    \centering
%    \begin{tabular}{c}
%    %\setlength{\tabcolsep}{0.1pt}
%    \subfigure[\Large\textwidth][\large Expt-$1$: $3$ Bernoulli-distributed arms (From Dr. Odalric's Draft).]
%    %with $r_{i_{{i}\neq {*}}}=0.07$ and $r^{*}=0.1$
%    {
%    		\pgfplotsset{
%		tick label style={font=\normalsize},
%		label style={font=\normalsize},
%		legend style={font=\normalsize},
%		ylabel style={yshift=12pt},
%		%legend style={legendshift=32pt},
%		}
%        \begin{tikzpicture}[scale=0.7]
%      	\begin{axis}[
%		xlabel={timestep},
%		ylabel={Cumulative Regret},
%		grid=major,
%        %clip mode=individual,grid,grid style={gray!30},
%        clip=true,
%        %clip mode=individual,grid,grid style={gray!30},
%  		legend style={at={(0.5,1.4)},anchor=north, legend columns=3} ]
%      	% UCB
%		
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_DUCB01.txt};
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_ETS04.txt};
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_ETS1E03.txt};
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_TS01.txt};
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_OTS01.txt};
%		\addplot table{results/NewExpt/Expt5/comp_subsampled_DTS01.txt};
%      	
%      	\legend{DUCB($\gamma=1-\frac{1}{4\sqrt{T}}$),ETSDAE4,ETSD1E3,TS,OTS,DTS($\gamma=1-\frac{1}{4\sqrt{T}})$}    
%      	\end{axis}
%      	\end{tikzpicture}
%  		\label{fig:2}
%    }
%    \end{tabular}
%    \caption{Cumulative regret for various bandit algorithms on a piecewise stochastic 3-armed bandit environment. }
%    \label{fig:karmed2}
%\end{figure}