### 2017-12-01 connBandits.tex

As defined in the previous section, an episode consist of a series of sequential interaction where as agent transitions from one state to another based on some intrinsic dynamics of the environment while collecting the rewards and choosing actions based on some action-selecting strategy. Now, for a single-step interaction, i.e., when the episode terminates after a single transition, the problem is captured by the multi-armed bandit (MAB) model. Infact the MAB model can be considered as a single looping state when the agent after taking an action, and observing the reward transitions back to the same state.