### 2017-12-01 Chapter1/studyBandit.tex

There are multiple reasons to study the interesting area of bandits. First of all, bandits are the cornerstone of understanding the general reinforcement learning area. Infact, bandits helps us to understand the idea of \textit{exploration-exploitation} dilemma which is the basis to build full, multi-state, general reinforcement learning ideas. Secondly, as stated in \citet{maillard2011apprentissage}, even $50$ years after \cite{robbins1952some} introduced the first idea of bandits, there are many interesting and fruitful areas where bandit concept can be extended in both practical and theoretical terms. Finally, there are several real-life industrial applications ranging from recommendation systems, game theory to anomaly detection where bandit applications have been found to perform exceptionally well. All of these forces us to delve deep into a systemic research of bandits.