### 2017-12-10 synopsis/objective.tex

The main objectives of the thesis are as follows:-
\begin{enumerate}
\item The first objective of this thesis is to study the area of stochastic multi-armed bandit (SMAB) and how to minimize cumulative regret in this setup. We intend to give strong gap-dependent and gap-independent regret guarantees in the SMAB setting. We also intend to provide algorithm in the SMAB setting that outperforms the current state-of-the-art algorithms in this setting.

\item The second objective of this thesis is to study the area of thresholding bandit problem (TBP) setting where the goal is to minimize the expected loss at the end of a fixed budget provided as input. We intend to provide strong guarantees with respect to expected loss and also propose algorithm that does not require any problem complexity as an input. We also intend to provide strong empirical evaluations of the algorithm proposed for the TBP setting.

%\item The third objective of this thesis is to study the area of piecewise stochastic bandit where again the goal is to minimize the cumulative regret. We intend to provide strong algorithms that can adapt to this environment and performs well empirically.  
\end{enumerate}