### 2018-01-08 abstract.tex

The thesis studies the following topics in the area of Reinforcement Learning: Multi-armed Bandits, Multi-armed bandits in stationary distribution with the goal of cumulative regret minimization and Thresholding bandits in pure exploration setting. The common underlying theme is the study of bandit theory and its application in various types of environments. In the first part of the thesis, we study the classic multi-armed bandit problem in stationary distribution, one of the first setting studied by the bandit community and which successively gave rise to several new directions in bandit theory. We propose a novel algorithm in this setting and compare both theoretically and empirically its performance against the available algorithms in this setting. Our proposed algorithm called Efficient-UCB-Variance (EUCBV) is the first arm-elimination algorithm which uses variance estimation yo eliminate arms as well as achieve an order optimal regret bound. Empirically, we show that EUCBV outperforms most of the state-of-the-art algorithms in the considered environments. In the next part, we study a very specific type of stochastic multi-armed bandit setup called the thresholding bandit problem and discuss extensively on its usage, available state-of-the-art algorithms on this setting and our solution to this problem. We propose the Augmented-UCB algorithm which again uses variance and mean estimation along with arm elimination technique to conduct exploration. We give theoretical guarantees on the expected loss of our algorithm and also analyze its performance against state-of-the-art algorithms in numerical simulations in multiple synthetic environments.