### 2019-05-30 Chapter1/contribution.tex
thesis initial draft

The main contributions of the thesis are as follows:-
\begin{enumerate}
\item We propose the MLTimer alogrithm that uses gate-sizing for reducing the leakage power consumption of a digital design. We propose a smart one-pass tool that can leverage the right optimization technique at the appropriate stage of the flow thereby improving design productivity. A key observation reported in MLTimer is that there exists significant correlation between the timing slacks of gates in the current iteration to the gate replacements in successive iterations. MLTimer leverages this observation to reduce the number of STA runs thereby reducing the overall time taken for optimization.

\item We propose the Karna algorithm which uses gate-sizing for reducing the information leakage via the power side-channel of a digital design. We show that each region in a given design leaks information differently. Thus, it is sufficient to optimize gates in the highly sensitive regions to reduce information leakage. Karna leverages this observation and optimizes gates in these sensitive regions to reduce the power side-channel vulnerability.

%\item We proposed a general framework of bandit algorithms that combines change-point detection algorithm with aggregation of expert strategies in order to define efficient pulling strategies in the context of the piecewise stochastic distributions. The algorithms that we proposed for the piecewise stochastic setting are actively adaptive algorithms which perform very similarly to the oracle algorithm which has access to the changepoints and suffers no additional delay in adapting to the changing environment. 
\end{enumerate}