### 2019-05-30 Chapter1/objective.tex

The main objectives of the thesis are as follows:-
\begin{enumerate}
\item The first objective of this thesis is to study the impact of gate-sizing on leakage power minimization. Our problem formulation takes into account the delay constraints of a given design. We also address the issues of scalability and, runtime.

\item The second objective of this thesis is to study the impact of gate-sizing as a possible countermeasure for power side-channel attacks. We propose a gate-sizing algorithm whose objective is to minimize information leakage via the power side-channel. We analyze the effectiveness of this algorithm by employing three different cipher designs. We also discuss the various factors affecting the convergence of the 
%area of thresholding bandit problem (TBP) setting where the goal is to minimize the expected loss at the end of a fixed budget provided as input. We intend to provide strong guarantees with respect to expected loss and also propose the algorithm that does not require any problem complexity as an input. We also intend to provide strong empirical evaluations of the algorithm proposed for the TBP setting.

%\item The third objective of this thesis is to study the area of piecewise stochastic bandit where again the goal is to minimize the cumulative regret. We intend to provide strong algorithms that can adapt to this environment and performs well empirically.  
\end{enumerate}